Accelerating Neural Network Training using Arbitrary Precision Approximating Matrix Multiplication Algorithms

### 2021-04-23 latex/ICPP21/paper.tex
add comments and new sections

\begin{abstract}
Matrix multiplication is one of the bottleneck computations for training the weights within deep neural networks.
To speed up the training phase, we propose to use faster algorithms for matrix multiplication known as Arbitrary Precision Approximating (APA) algorithms.
APA algorithms perform asymptotically fewer arithmetic operations than the classical algorithm, but they compute an approximate result with an error that can be made arbitrarily small in exact arithmetic.
Practical APA algorithms provide significant reduction in computation time and still provide enough accuracy for many applications like neural network training.
We demonstrate that APA algorithms can be efficiently implemented and parallelized for multicore CPUs to obtain up to 28\% and 21\% speedups over the fastest implementation of the classical algorithm using one core and 12 cores, respectively.
Furthermore, using these algorithms to train a Multi-Layer Perceptron (MLP) network yields no significant reduction in the training or testing error.
Our performance results on a large MLP network show overall sequential and multithreaded performance improvements of up to 25\% and 13\%, respectively.
\end{abstract}

@@

Our contribution is the use of APA matrix multiplication algorithms to address this problem.
In particular we
\begin{enumerate}
	\item curate a collection of both well-established and recently discovered practical APA algorithms;
	\item extend the framework of \cite{BB15} to generate efficient multithreaded code for all of them, achieving up to 21\% performance improvement over the best parallel classical implementation,
	\item demonstrate the robustness of learning accuracy to approximate matrix multiplications; and
	\item present multithreaded performance improvements of a synthetic MLP of up to 13\% over the use of classical matrix multiplication.
\end{enumerate}

@@

@@

>>>>

\begin{abstract}
Matrix multiplication is one of the bottleneck computations for training the weights within deep neural networks.
To speed up the training phase, we propose to use faster algorithms for matrix multiplication known as Arbitrary Precision Approximating (APA) algorithms.
APA algorithms perform asymptotically fewer arithmetic operations than the classical algorithm, but they compute an approximate result with an error that can be made arbitrarily small in exact arithmetic.
Practical APA algorithms provide significant reduction in computation time and still provide enough accuracy for many applications like neural network training.
We demonstrate that APA algorithms can be efficiently implemented and parallelized for multicore CPUs to obtain up to 28\% and 21\% speedups over the fastest implementation of the classical algorithm using one core and 12 cores, respectively.
Furthermore, using these algorithms to train a Multi-Layer Perceptron (MLP) network yields no significant reduction in the training or testing error.
Our performance results on a large MLP network show overall sequential and multithreaded performance improvements of up to 25\% and 13\%, respectively.
\RZ{We also demonstrate that APA algorithms are efficiently implemented on a more complicated network VGG-11?}
\end{abstract}

@@

Our contribution is the use of APA matrix multiplication algorithms to address this problem.
In particular we
\begin{enumerate}
	\item curate a collection of both well-established and recently discovered practical APA algorithms;
	\item extend the framework of \cite{BB15} to generate efficient multithreaded code for all of them, achieving up to 21\% performance improvement over the best parallel classical implementation;
	\item demonstrate the robustness of learning accuracy to approximate matrix multiplications;
	\item present multithreaded performance improvements of a synthetic MLP of up to 13\% over the use of classical matrix multiplication; and
	\item \RZ{present performance improvements of VGG-11? network of ?\% over the classical matrix multiplication.}
\end{enumerate}

@@

\subsection{Skewed Matrix Multiplication}
When computing square matrix multiplications, we have demonstrated that $\dims{4,4,2}$ and $\dims{4,4,4}$ algorithms have the best performance.
However, algorithms with more skewed aspect ratios will likely perform better for problems with matching skewed matrix multiplication dimensions.
We note that an algorithm for dimensions $\dims{m,n,k}$ can be translated into an algorithm for $\dims{n,m,k}$ and any other reordering of the dimensions \cite{BB15}.
\RZ{transposing row major matrix problem?}
\RZ{Best algorithm for each shape of matrices and its figure}

@@

\section{Application of APA algorithms in VGG-11}
\subsection{VGG-11 Network}
\RZ{description of the model of VGG-11 and VGG-19?}

% \subsection{Accuracy}

\subsection{Performance}
\RZ{VGG-11 fully connected layer: BS x 25K x 4K / BS x 4K x 4K / BS x 4K x 1K? with plot}
